Current and future constraints on primordial magnetic fields

### 2015-11-10 pmf.tex
intro

Precision measurements of cosmic microwave background (CMB) polarization open up new tests for extensions to the  standard cosmological model. 
One such extension would be the existence of a primordial magnetic field (PMF).

@@

In addition to the \planck{} data, we use all recent CMB polarization measurements that are publicly available. 
First, we include TE and EE spectrum measurements\citep{crites15} for $\ell \in [500,3000]$ and BB bandpowers covering $\ell \in [500,2000]$ \citep{keisler15} from SPTpol. 
We also add the BB bandpowers from \pb{} that cover the multipoles from 500 to 2500 \citep{pb-bb15}. 
Both the \sptpol{} and \pb{} bandpowers primary constrain the vector mode of the PMF given the angular scales measured. 
Finally, we include the latest BICEP2 and Keck Array  joint analysis \citep{bicepkeck}. 
This latter dataset also places limits on the tensor mode of the PMF due to its coverage of lower multipoles. 
In all chains, we marginalize over the recommended foreground models for each data set. 
We do not however require these consistency between these foreground models since the data don't have identical flux cuts and so on. 
In the case of an upper limit this treatment is conservative since weaker foreground constraints or residual point source power would only add power and weaken the limit on the PMF.

@@

The PMF power spectrum template is computed by a modified code CMBACT\footnote{\url{http://www.sfu.ca/~levon/cmbact.html}} with parameters $B_{\rm{1Mpc}}$=2.5 \rm{nG}, $n_s$=-2.9 and $a_{\nu}/a_{\rm{PMF}}$=$10^9$. 
We can translate the amplitude constraint on this template to a constraint rms magnetic field strength on 1\,Mpc scales , \bpmf, using the expected scaling:
\be
\apmf \propto \bpmf^4.
\ee
An unfortunate consequence of this steep scaling is that substantial reductions in the observationally allowed power lead to only modestly better limits on the magnetic field strength. 
However, if a PMF is detected, this same steep scaling would allow excellent constraints on its properties.

>>>>

Measurements of the cosmic microwave background (CMB) temperature anisotropy have provided some of the most powerful tests of cosmology. 
With a number of experiments beginning to measure  ``B-modes" in CMB polarization for the first time \citep{XXXX}, we are entering a new era for CMB observations. 
Precision measurements of cosmic microwave background (CMB) polarization promise new tests of the standard cosmological model. 
The best known of these tests are the searches for inflationary gravitational waves in B-modes at large angular scales \citep{XXX} and plans to measure the sum of the neutrino masses through the lensing B-modes on small angular scales \citep{XXX}.

CMB B-mode measurements will constrain more exotic models as well, such as the possible existence of cosmic birefringence \citep{carroll98,lue99} or a primordial magnetic field (PMF) \citep{kosowsky96, seshadri01}.  
In this work, we present new upper limits from current CMB polarization data on the possibility of a PMF or parity-violating physics. 
Both effects lead to birefringence and the rotation of E-modes into B-modes. 
By equating the magnitude of the resulting B-modes, parity-violating processes can be translated into an equivalent PMF strength. 
We will therefor simply quote effective limits on a PMF.

Parity-violating processes

However, CMB polarization data are expected to improve constraints on extensions

of precision measurements of CMB polarization. 
CMB polarization opens up

we are beginning a new era of precision measurements of CMB polarization.

These measurements are now being extended to CMB polarization

one of the main pillars of the standard cosmolog
Precision measurements of cosmic microwave background (CMB) polarization open up new tests for extensions to the  standard cosmological model. 
Such extensions include the

@@

In addition to the \planck{} data, we use all recent CMB polarization measurements that are publicly available. 
First, we include TE and EE spectrum measurements\citep{crites15} for $\ell \in [500,3000]$ and BB bandpowers covering $\ell \in [500,2000]$ \citep{keisler15} from SPTpol. 
We also add the BB bandpowers from \pb{} that cover the multipoles from 500 to 2500 \citep{polarbear14b}. 
Both the \sptpol{} and \pb{} bandpowers primary constrain the vector mode of the PMF given the angular scales measured. 
Finally, we include the latest BICEP2 and Keck Array  joint analysis \citep{bicepkeck15}. 
This latter dataset also places limits on the tensor mode of the PMF due to its coverage of lower multipoles. 
In all chains, we marginalize over the recommended foreground models for each data set. 
We do not however require these consistency between these foreground models since the data don't have identical flux cuts and so on. 
In the case of an upper limit this treatment is conservative since weaker foreground constraints or residual point source power would only add power and weaken the limit on the PMF.

@@

The PMF power spectrum template is computed by a modified code CMBACT\footnote{\url{http://www.sfu.ca/~levon/cmbact.html}} \citep{pogosian99} with parameters $B_{\rm{1Mpc}}$=2.5 \rm{nG}, $n_s$=-2.9 and $a_{\nu}/a_{\rm{PMF}}$=$10^9$. 
We can translate the amplitude constraint on this template to a constraint rms magnetic field strength on 1\,Mpc scales , \bpmf, using the expected scaling:
\be
\apmf \propto \bpmf^4.
\ee
An unfortunate consequence of this steep scaling is that substantial reductions in the observationally allowed power lead to only modestly better limits on the magnetic field strength. 
However, if a PMF is detected, this same steep scaling would allow excellent constraints on its properties.
